---
title: "Homework Week 6"
author: "Andrew"
date: "October 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question 9.1

```{r}
library(ggplot2)
library(dplyr)
library(DAAG)
library(corrplot)
library(broom)
library(factoextra)
```

```{r}
##Reading the crime data files into R and saving as crime
crime<- read.delim("uscrime.txt", sep = "\t", header = TRUE)
```

#Explanation of each variable:
#Variable	 	Description
#M		percentage of males aged 14-24 in total state population
#So		indicator variable for a southern state
#Ed		mean years of schooling of the population aged 25 years or over
#Po1		per capita expenditure on police protection in 1960
#Po2		per capita expenditure on police protection in 1959
#LF		labour force participation rate of civilian urban males in the age-group 14-24
#M.F		number of males per 100 females
#Pop		state population in 1960 in hundred thousands
#NW		percentage of nonwhites in the population
#U1		unemployment rate of urban males 14-24
#U2		unemployment rate of urban males 35-39
#Wealth		wealth: median value of transferable assets or family income
#Ineq		income inequality: percentage of families earning below half the median income
#Prob		probability of imprisonment: ratio of number of commitments to number of offenses
#Time		average time in months served by offenders in state prisons before their first release
#Crime		crime rate: number of offenses per 100,000 population in 1960

```{r}
##inspecting the first five rows of crime, "Crime" is crimes per 100K and the data I will be looking to predict
head(crime)
```
```{r}
##Creating a dataframe of the predictor variables, i.e. a dataframe that does not include Crime
crime_two<- crime[,1:15]
head(crime_two)
```
```{r}
##Creating a dataframe of a given city's data to predict the crime rate using the forthcoming regression models
predict_city<- matrix(c(14,0,10,12,15.5,0.64,94,150,1.1,0.12,3.6,3200,20.1,0.04,39), nrow = 1, ncol = 15)
predict_city<- as.data.frame(predict_city)
names<- colnames(crime[1:15])
colnames(predict_city)<- names
rownames(predict_city)[1]<- "city"
```

```{r}
##Principal component analysis on the crime data
pr_test<- prcomp(crime_two, scale = TRUE)
summary(pr_test)
```
```{r}
##Correlation between principal components
cor(pr_test$x)
```
##The correlation bewtween the model's 15 principal components is essentially zero

```{r}
##Chart of proportion of variance explained by each PC
fviz_eig(pr_test)
```

##Given that the largest percentage of explained variances is gained around 5 dimensions, I will use the first five principal components in my first regression model.  Afterwards I will create a model using only the first four principal components to compare results.

```{r}
##Matrix containing the eigenvectors for each PC
pr_test$rotation
```



```{r}
##Eigenvalue for each PC, percent variance captured for each, and the cumulative variance in percent captured
get_eigenvalue(pr_test)
```
```{r}
##Getting first 5 principal components which explain capture 86.31% of the variance
(PC<- pr_test$x[,1:5])
```
```{r}
##Regression model using first five principal components, first combining PC's and response variable
crime_pc<- cbind(PC, crime[,16])
colnames(crime_pc)[6]<- "Crime"
head(crime_pc)
```
```{r}
##Creating regression model using newly combined data
pca_model<- lm(Crime ~ . , data = as.data.frame(crime_pc))
summary(pca_model)
```

##The regression model using the first five principal components returns an adjusted r-squared value of 0.6019
```{r}
##Reversing PCA, matrix multiplication of the model coefficients by the transposed eigenvectors for the first five PC's
pca_rev<- pca_model$coefficients[2:6] %*% t(pr_test$rotation[,1:5])
head(pca_rev)
```

```{r}
##Unscaling coefficients and intercept (scale under the PCA = standard deviations, center = mean)
alpha_unscaled<- pca_rev/pr_test$scale
intcpt<- pca_model$coefficients[1] - sum(pca_rev*pr_test$center / pr_test$scale)
```


```{r}
##Using unscaled coefficients to predict the crime rate in the given city and adding the unscaled intercept
(CR<- sum(alpha_unscaled * predict_city) + intcpt)
```

##The predicted Crime rate for the test city using the first five principal components is 1388


##Re-running the regression using only the first four principal components

```{r}
##Getting first 4 principal components which explain capture 79.92% of the variance
PC4<- pr_test$x[,1:4]
```

```{r}
##Regression model using first four principal components, first combining PC's and response variable
crime_pc4<- cbind(PC4, crime[,16])
colnames(crime_pc4)[5]<- "Crime"
head(crime_pc4)
```
```{r}
##Creating regression model using newly combined data
pca_model4<- lm(Crime ~ . , data = as.data.frame(crime_pc4))
summary(pca_model4)
```

##The adjusted r-square for the regression model using only the first four principal components falls from 0.60 in the first model to 0.24

```{r}
##Reversing PCA into unscaled coefficients
pca_rev4<- pca_model4$coefficients[2:5] %*% t(pr_test$rotation[,1:4])
head(pca_rev4)
```
```{r}
##Unscaling coefficients and intercept
alpha_unscaled4<- pca_rev4/pr_test$scale
intcpt4<- pca_model4$coefficients[1] - sum(pca_rev4*pr_test$center / pr_test$scale)
```


```{r}
##Using unscaled coefficients to predict the crime rate in the given city
(CR4<- sum(alpha_unscaled4 * predict_city) + intcpt4)
```

##The predicted crime rate using only the first four principal components is 1112.678, which is slightly less than the amount predicted by the model using five principal components.


##Using both models based on either the first five our four principal components result in reasonable crime rates.  Looking at the Q-Q plot shows that the estimates fall near the middle of the distribution.  Compared to the original models from last week's homework, the adjusted r-squared values are comparable; the model using five principal components has an adjusted r-square of 0.60, while the model using four principal components the adjusted r-square falls to 0.2433.












