---
title: "Homework 7_2 Clean"
author: "Andrew"
date: "September 21, 2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(xts)
library(lubridate)
library(smooth)
library(ggplot2)
library(reshape2)
library(Rcpp)
library(guf)
library(dplyr)
```

#Question 7.2

#Using the 20 years of daily high temperature data for Atlanta (July through October) from Question 6.2 (file temps.txt), build and use an exponential smoothing model to help make a judgment of whether the unofficial end of summer has gotten later over the 20 years.  (Part of the point of this assignment is for you to think about how you might use exponential smoothing to answer this question. Feel free to combine it with other models if you'd like to. There's certainly more than one reasonable approach.) 

#Note: in R, you can use either HoltWinters (simpler to use) or the smooth package's es function (harder to use, but more general).  If you use es, the Holt-Winters model uses model="AAM" in the function call (the first and second constants are used "A"dditively, and the third (seasonality) is used "M"ultiplicatively; the documentation doesn't make that clear). 


```{r}
##Reading the temperature data files into R and saving as temps
temps<- read.delim("temps.txt", sep = "\t", header = TRUE)
```

```{r}
##Viewing the first few rows of temps
head(temps)
```

```{r}
##Summary statistics of temps
summary(temps)
```

```{r}
##Adding daily average column to data
temps_avg<- temps
temps_avg$avg<- rowMeans(temps[,-1])
```

```{r}
##Ploting average daily temperature over the 1996 to 2015 time period
plot(temps_avg$avg, type = "l", xaxt = "n", xlab = "Date", ylab = "Temperature", main = "Average Daily Temp. 1996-2015", col = "red")
axis(1, at = 1:123, labels = temps$DAY, las = 2, cex.axis = 0.75)
```


```{r}
##Ploting yearly daily temperature over the 1996 to 2015 time period
plot(temps_avg$X1996, type = "l", xaxt = "n", xlab = "Date", ylab = "Temperature", main = "Temperature by Year 1996-2015", col = "red")
lines(temps_avg[,3], col = "blue")
lines(temps_avg[,4], col = "green")
lines(temps_avg[,5], col = "gray")
lines(temps_avg[,6], col = "yellow")
lines(temps_avg[,7], col = "purple")
lines(temps_avg[,8], col = "orange")
lines(temps_avg[,9], col = "black")
lines(temps_avg[,10], col = "steelblue")
lines(temps_avg[,11], col = "magenta")
lines(temps_avg[,12], col = "cyan")
lines(temps_avg[,13], col = "pink")
lines(temps_avg[,14], col = "brown")
lines(temps_avg[,15], col = "tan")
lines(temps_avg[,16], col = "coral")
lines(temps_avg[,17], col = "cadetblue")
lines(temps_avg[,18], col = "blueviolet")
lines(temps_avg[,19], col = "cornsilk")
lines(temps_avg[,20], col = "darkgray")
lines(temps_avg[,21], col = "burlywood")
lines(temps_avg[,22], col = "midnightblue", lwd = 4)
axis(1, at = 1:123, labels = temps$DAY, las = 2, cex.axis = 0.75)
```

```{r}
##Melt dataset into long format
temps_long<- melt(temps, id = "DAY")
```

```{r}
##Manipulating the data labels
temps_long$variable<- as.character(temps_long$variable)
temps_long$variable<- right(temps_long$variable,4)
colnames(temps_long)[2]<- "year"
colnames(temps_long)[3]<- "temp"
```

```{r}
##Merging the day column and year column together, converting to a date based object, relocating columns for easier reading
temps_long$date<- as.Date(paste(temps_long$DAY,temps_long$year,sep = "-"), "%d-%b-%Y")
temps_long<- temps_long %>% 
  relocate(date, .before = DAY) %>%
  relocate(temp, .after = date)
```

```{r}
##Removing redundant columns
temps_long_clean<- temps_long[,1:2]
```


```{r}
##Convert temps_long into a time-series object (ts)
temps_ts<- ts(temps_long_clean$temp, start = 1996, frequency = 123)
head(temps_ts)
tail(temps_ts)
```

```{r}
##Decomposition of the temperature time series
dec_ts<- decompose(temps_ts)
plot(dec_ts)
```

#The above decomposition of the temperature time series shows that, on average, there is little trend in the data, but there is a steady seasonal pattern

```{r}
##Running Holt Winters model on time series of temperature data, addititve seasonality
hw<- HoltWinters(temps_ts)
summary(hw)
hw$alpha
hw$beta
hw$gamma
hw$seasonal
hw$SSE
```
#The above Holt Winters model, using additive seasonality, generated a model with 66244.25 SSE and an alpha coefficient of 0.6610618.

```{r}
##Plotting the components of the Holt Winters model
plot(hw$fitted)
```


```{r}
##Plot of the fitted values from the additive Holt Winters model vs. actual data
plot(hw)
```

#The above plot of the modeled data (red line) against the actual temperature data (black line) show that the modeled data appears to 'fit' the actual data better in later years
```{r}
##Running Holt Winters model on time series of temperature data, multiplicative seasonality
hw_m<- HoltWinters(temps_ts, seasonal = "multiplicative")
summary(hw)
hw$alpha
hw$beta
hw$gamma
hw$seasonal
hw$SSE
```
#I re-ran the Holt Winters model, but with multiplicative seasonaility this time instead of additive.  Alpha and SSE remain the same.
```{r}
##Plotting components of the Holt Winters model
plot(hw_m$fitted)
```

```{r}
##Plotting the Holt Winters fitted values vs. actual temperature data
plot(hw_m)
```

```{r}
##Converting the data into a matrix (similar to the original temps dataset)
fitted_mtx<- matrix(hw$fitted[,1], nrow = 123)
season_mtx<- matrix(hw$fitted[,4], nrow = 123)
fitted_mtx_m<- matrix(hw_m$fitted[,1], nrow = 123)
season_mtx_m<- matrix(hw_m$fitted[,4], nrow = 123)
```




#Using excel, I will now perform change detection using CUSUM to determine if the unoffical end of summer has gotten later over the 20year period of the dataset.
